---
title: "1. R Basics"
output: rmarkdown::github_document
---

1. Install and Load Packages

Install packages: once you successfully install the packages, you don't need to install them again in the same environment.

```{r}
# An R package is a collection of functions, data, and documentation that extends the capabilities of base R
# Tidyverse is one of most popular sets of packages in R for data analysis
install.packages("tidyverse")
install.packages("tidytext")
```

Load packages: you need to load the packages when you want to use them

```{r}
library(tidyverse)
library(tidytext)
```


2. Basic Operations in R

```{r}

# You can do all kinds of basic computation in R
# More on basic operation in R:
# https://www.cyclismo.org/tutorial/R/basicOps.html

1 + 1 # add
```

```{r}

# In R, we use <- to assign a value to a variable
# More on variable in R:
# https://www.tutorialspoint.com/r/r_variables.htm

first_variable <- 500 + 700

print(first_variable)
```

3. Data Frame basics

A data frame is a table like structure in which each column contains values of one variable and each row contains one set of values from each column. 

Tidyverse packages offers many functions on handling data in data frame format.

```{r}

# load a built-in data
data("mtcars")

# to see the first few rows of the data
head(mtcars)
```
```{r}

# to see the whole dataset in a new window in R studio
# don't use this function if the data set is too large
View(mtcars)

# more on data frames basics
# https://www.tutorialspoint.com/r/r_data_frames.htm
```

